\ppse{} - A Peer-to-Peer Support for Multiplayer Games

### 2008-08-12 trunk/textos/p2pse-sbgames2008/ppsg.tex
blinded

\newcommand{\ppse}{Proj48534}

@@

% %% Author and Affiliation (multiple authors). Use: and between authors
% \author{Felipe J. Vilanova, Carlos Eduardo B. Bezerra , Marcos R. Crippa, Fábio R. Cecin, Cláudio F. R. Geyer\\
%           Universidade Federal do Rio Grande do Sul\\
%           Bento Gonçalves, 9500\\
%           Porto Alegre, RS, Brazil\\
%           +55 (51) 3316 0000
% }
% \contactinfo{\{fjvilanova, carlos.bezerra, mrcrippa, fcecin, geyer\}inf.ufrgs.br}

@@

In this paper a hybrid model is proposed, combining the security and consistency of the client-server model with the scalability and flexibility of the distributed model. In sections \ref{sec:c-smodel} and \ref{sec:p2pmodel} we discuss about Client-server and Peer-to-Peer models. The \ppse{} %(acronym in portuguese for ``P2P for Entertainment Software'') 
architecture, is explained in section \ref{sec:p2pse}. Section \ref{sec:simulations} presents the simulations we made to compare client-server and peer-to-peer models. Section \ref{sec:conclusion} presents conclusions on the topic.

@@

The client-server paradigm is undoubtedly the most currently used in the implementation of Internet real-time multiplayer games. Commercial examples of these games include Doom, Quake \cite{quake} and Counter-Strike \cite{counterstrike}, all them 3D real time action games. In any client-server game, massively multiplayer or not, the players interact with the environment and with other players through a client program. Each client has a network connection only with the server. The server is responsible for receiving the information from each client and passing updates to the other clients. There are several different ways the communication protocol between the client and the server can be implemented: some of them provide more security, while others provide efficiency.

Regardless of the technique used, the server is responsible for maintaining an updated state of the game. Since the game is a real time simulation, it is necessary for the state of the game to be frequently recalculated. In general, the server is configured to update its state in a fixed and relatively small frequency, and to periodically send update messages to all clients. Due to that, the server must have great processing power (CPU and, indirectly, memory) and enough bandwidth (network).

@@

There are other proposals that do not address the cheating problem, such as Hydra \cite{chan2007hmm}, that focuses on guaranteeing the consistency in the messages committed when nodes fail; and Mediator \cite{fan2007mdf}, which address the scalability and performance problems by proposing a solution based on a super-peer network and a reward scheme: peers that contribute more can use more resources.

@@

Depending on the game, the resulting conflicts from a lack of centralized timestamping of movement and similar events, whenever they ensue, can be either handled by each peer individually or left for the super-peer to detect it and issue a special correction message later, as the BZFlag protocol does \cite{pellegrino2003bra}. Local corrections are an option if each peer is authoritative over its own avatar's position and if it broadcasts position updates for its own avatar. For instance, if a peer's local avatar overlaps with a remote peer's, the local avatar can push its own avatar out of that position. The remote avatar will do the same. The only issue is ensuring that the correction step won't make both peers try to resolve the situation by unstucking their avatars to the same spot again. This local conflict resolution is what we are currently employing in our capture-the-flag action prototype, Hoverkill \cite{hoverkill}.

@@

In this project we tried a somewhat unique approach. Since most events would be still centralized at the super-peer and most action games have one-per-client avatar position updates as their consistency, bandwidth and latency bottlenecks, we just left peers with total control over their positions and with the responsibility for sending them directly to other peers in a broadcast fashion. This optimizes for consistency and latency, and doesn't create a bandwidth bottleneck at the super-peer. On top of that, we tried AI-based detection of unusual patters over streams of position updates coming from each peer. The main motivation was to detect blatant speed-cheating being performed by players while other players (without admin powers, nevertheless) quit the game in frustration. It was not our goal to prevent subtle position manipulation, since we wanted to avoid false positives more than we wanted to prevent any form of position cheating.

We have trained, using back-propagation, an Artificial Neural Network (ANN) over traces of gameplay generated from Hoverkill play sessions, which is our actual game prototype. We obtained some positive initial results \cite{gaspareto2008nna},  but we also found that a successful ANN detection engine is tied to many dynamic parameters, including specific game rules and maps, general behavior of game players (if it changes over time, the network must be re-trained), quality of the traces (they must cover all the detectable patterns and with enough variation), and others. The results put some merit on the general idea that using an automatic detection engine, instead of human attention, to discourage illegal position manipulation during gameplay could enable more simple and efficient but vulnerable protocols such as ours to still succeed in a real MMOG deployment scenario.

@@

It is important to notice, however, that the time a player usually spends playing action games, such as first-person shooters, is much longer than the time they keep talking in the game chat room. Therefore, the average time spent by each player in the social space should be much shorter than the time spent on the action spaces in the simulation. Also, the social space will most likely have much less need for frequent updates, allowing further decrease of the bandwidth utilization by the \ppse{} server. Anyway, those parameters were chosen considering a pessimistic scenario, where the social space is as interactive as the action space, and the players spend more time socializing than it usually occurs in this kind of games.

@@

We are currently integrating the \ppse{} architecture with our game prototype, Hoverkill. This game is a client-server capture-the-flag tank game, and its network system is being replaced by our architecture. The final goal is to be able to test the game in client-server and peer-to-peer modes, in order to base our conclusions in real results rather than simulations..

>>>>

\newcommand{\ppse}{Proj48534}
\newcommand{\hoverkill}{HK48534}

@@

\author{JEMS Manuscript Number 48534}
\contactinfo{48534@JEMS}
% %% Author and Affiliation (multiple authors). Use: and between authors
% \author{Felipe J. Vilanova, Carlos Eduardo B. Bezerra , Marcos R. Crippa, Fábio R. Cecin, Cláudio F. R. Geyer\\
%           Universidade Federal do Rio Grande do Sul\\
%           Bento Gonçalves, 9500\\
%           Porto Alegre, RS, Brazil\\
%           +55 (51) 3316 0000
% }
% \contactinfo{\{fjvilanova, carlos.bezerra, mrcrippa, fcecin, geyer\}inf.ufrgs.br}

@@

In this paper a hybrid model is proposed, combining the security and consistency of the client-server model with the scalability and flexibility of the distributed model. In sections \ref{sec:c-smodel} and \ref{sec:p2pmodel} we discuss the client-server and peer-to-peer models. The \ppse{} %(acronym in portuguese for ``P2P for Entertainment Software'') 
architecture is explained in section \ref{sec:p2pse}. Section \ref{sec:simulations} presents the simulations we made to compare client-server and peer-to-peer models. Section \ref{sec:conclusion} presents conclusions on the topic.

@@

The client-server paradigm is undoubtedly the most currently used in the implementation of Internet real-time multiplayer games. Commercial examples of these games include Doom \cite{doom}, Quake \cite{quake} and Counter-Strike \cite{counterstrike}, all them 3D real time action games. In any client-server game, massively multiplayer or not, the players interact with the environment and with other players through a client program. Each client has a network connection only with the server. The server is responsible for receiving the information from each client and passing updates to the other clients. There are several different ways the communication protocol between the client and the server can be implemented: some of them provide more security, while others provide efficiency.

Regardless of the technique used, the server is responsible for maintaining the game state up-to-date. Since the game is a real time simulation, it is necessary for the state of the game to be frequently recalculated. In general, the server is configured to update its state in a fixed and relatively small frequency, and to periodically send update messages to all clients. Due to that, the server must have great processing power (CPU and, indirectly, memory) and enough bandwidth (network).

@@

There are other proposals that do not address the cheating problem, such as Hydra \cite{chan2007hmm}, that focuses on guaranteeing the consistency in the messages committed when nodes fail; and Mediator \cite{fan2007mdf}, which addresses the scalability and performance problems by proposing a solution based on a super-peer network and a reward scheme: peers that contribute more can use more resources.

@@

Depending on the game, the resulting conflicts from a lack of centralized timestamping of movement and similar events, whenever they ensue, can be either handled by each peer individually or left for the super-peer to detect it and issue a special correction message later, as the BZFlag protocol does \cite{pellegrino2003bra}. Local corrections are an option if each peer is authoritative over its own avatar's position and if it broadcasts position updates for its own avatar. For instance, if a peer's local avatar overlaps with a remote peer's, the local avatar can push its own avatar out of that position. The remote avatar will do the same. The only issue is ensuring that the correction step won't make both peers try to resolve the situation by unstucking their avatars to the same spot again. This local conflict resolution is what we are currently employing in our capture-the-flag action prototype, \hoverkill\ (reference blinded).%Hoverkill \cite{hoverkill}.

@@

In this project we tried a somewhat unique approach. Since most events would be still centralized at the super-peer and most action games have one-per-client avatar position updates as their consistency, bandwidth and latency bottlenecks, we just left peers with total control over their positions and with the responsibility for sending them directly to other peers in a broadcast fashion. This optimizes for consistency and latency, and doesn't create a bandwidth bottleneck at the super-peer. On top of that, we tried AI-based detection of unusual patterns over streams of position updates coming from each peer. The main motivation was to detect blatant speed-cheating being performed by players while other players (without admin powers, nevertheless) quit the game in frustration. It was not our goal to prevent subtle position manipulation, since we wanted to avoid false positives more than we wanted to prevent any form of position cheating.

We have trained, using back-propagation, an Artificial Neural Network (ANN) over traces of gameplay generated from \hoverkill\ play sessions, which is our actual game prototype. We obtained some positive initial results [reference blinded]%\cite{gaspareto2008nna}
,  but we also found that a successful ANN detection engine is tied to many dynamic parameters, including specific game rules and maps, general behavior of game players (if it changes over time, the network must be re-trained), quality of the traces (they must cover all the detectable patterns and with enough variation), and others. The results put some merit on the general idea that using an automatic detection engine, instead of human attention, to discourage illegal position manipulation during gameplay could enable more simple and efficient but vulnerable protocols such as ours to still succeed in a real MMOG deployment scenario.

@@

It is important to notice, however, that the time a player usually spends playing action games, such as first-person shooters, is much longer than the time they keep talking in the game chat room. Therefore, the average time spent by each player in the social space should be much shorter than the time spent on the action spaces in the simulation. Also, the social space will most likely have much less need for frequent updates, allowing further decrease of the bandwidth utilization by the \ppse{} servers. Anyway, those parameters were chosen considering a pessimistic scenario, where the social space is as interactive as the action space, and the players spend more time socializing than it usually occurs in this kind of games.

@@

We are currently integrating the \ppse{} architecture with our game prototype, \hoverkill{}. This game is a client-server capture-the-flag tank game, and its network system is being replaced by our architecture. The final goal is to be able to test the game in client-server and peer-to-peer modes, in order to base our conclusions in real results rather than simulations.
